---
title: "Reproducible_Research_Course_Project_2"
author: "J Go"
date: "June 14, 2016"
output: html_document
---

### Title: An Exploration of the NOAA Storm Database to Determine Health and 
###        Economic Effects of Severe Weather Events in the United States

### Synopsis

```
"Storms and other severe weather events can cause both public health and 
economic problems for communities and municipalities. Many severe events can 
result in fatalities, injuries, and property damage, and preventing such 
outcomes to the extent possible is a key concern."  [course assignment page]

In this report the NOAA Storm Database is explored to answer two basic questions 
about severe weather events. 

1. Across the United States, which types of events (as indicated in the 
ğ™´ğš…ğšƒğšˆğ™¿ğ™´ variable) are most harmful with respect to population health?

2.  Across the United States, which types of events have the greatest economic 
consequences?

This report is in a "literate statistical programming" format with the intention 
of making the entire analysis readily reproducible. 

Possible audiences may include: government or municipal managers responsible for 
preparing for severe weather events who need to prioritize resources for 
different types of events, and fellow students and practitioners of data science 
who may find the methods of analysis and documentation of interest. 

Disclaimer: Please note that the analysis is for educational purposes only.  It 
does not make any determination or recommendation for any specific course of 
action or measure to be taken.  Further the analysis does not constitute an 
endorsement or validation of any of the methods, software, hardware, or results 
used or obtained herein.
```

### Data Processing

There should be a section titled Data Processing which describes (in words and code) how the data were loaded into R and processed for analysis. In particular, your analysis must start from the raw CSV file containing the data. You cannot do any preprocessing outside the document. If preprocessing is time-consuming you may consider using the ğšŒğšŠğšŒğš‘ğš = ğšƒğšğš„ğ™´ option for certain code chunks.

There should be a section titled Results in which your results are presented.

You may have other sections in your analysis, but Data Processing and Results are required.
The analysis document must have at least one figure containing a plot.
Your analysis must have no more than three figures. Figures may have multiple plots in them (i.e. panel plots), but there cannot be more than three figures total.
You must show all your code for the work in your analysis document. This may make the document a bit verbose, but that is okay. In general, you should ensure that ğšğšŒğš‘ğš˜ = ğšƒğšğš„ğ™´ for every code chunk (this is the default setting in knitr).
```

```
### Loading and Processing the Raw Data

The U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database 
tracks characteristics of major storms and weather events in the United States, 
including when and where they occur, as well as estimates of any fatalities, 
injuries, and property damage.

First, data fom the NOAA storm database is read in from the course web site.  
The raw text is contained in a .bz2 compressed file.  The data fields are 
comma delimited.

### check for repdata-data-StormData.csv.bz2 file, download into data dir if necessary
```{r check_existence_file_repdata-data-StormData.csv.bz2}
data_file <- paste0(file.path(".", "data", "repdata-data-StormData.csv.bz2"))
if(!file.exists(data_file)) {
    file1Url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
    download.file(file1Url, destfile = "./data/repdata-data-StormData.csv.bz2", method = "curl")
    dateDownloaded <- date()
} else {
    cat('data file: [', data_file, '] was downloaded in a previous run \n')
    cat('on [', dateDownloaded, '] \n')
}
```

```{r read storm data file}
# storm <- read.csv(bzfile("./data/repdata-data-StormData.csv.bz2")) 

# if (!exists("storm")) {  # this not working
# } else {
#   cat('storm data was read in during a previous run\n')
# }
str(storm)
summary(storm)
```
